---
title: 'Statistics: Correlation Analysis'
author: "Scott Schumacker"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project will show an example correlation analysis on female height and weight from the 'women' data set on R.

***Loading Packages***
```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(tidyr)
```

## Creating a data set
For this example of a correlation analysis, we can use the women data set. Here is a preview of the data
```{r}
# Preview data
head(women)
```

## Visualizing the trend of the data.
After obtaining our data set, we can visually inspect the relationship between the two variables:
```{r}
women %>% 
  ggplot(aes(x = height, y = weight)) +
  geom_point(size = 3, alpha = 0.8) +
  theme_classic()
```

We can see that there is a clear linear trend between these two variables.

## Statistical Analysis
Now that we know there is most likely a strong linear correlation between the two variables we can compute the correlation coefficient and run Pearson's product-moment correlation test.

Before we run the test we can create our hypothesis:
H0: The true correlation between the variables is equal to 0
Ha: The true correlation between the variables in not equal to 0

We can run this test using an alpha of 0.05.
```{r}
# Calculating correlation coefficient
cor(x = women$height, y = women$weight)
```

```{r}
# Running Pearson's product-moment correlation
cor.test(x = women$height, y = women$weight)
```

After running Pearson's product-moment correlation test, we can see that the correlation estimate is 0.995 with a 95% CI (0.986, 0.999) with a p-value < 0.01 (p-value = 1.091e-14). Since our p-value is less than our alpha we can reject the null hypothesis and assume that the true correlation between these variables is not 0.